The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-03-24 Paper/jaccardbaetica.tex
discussion britannia

The paper is structured as follows: the next section introduces the historical context. Section two displays the dataset and the methods used for the analysis. Section three presents the results and the last section shows the discussion and the main conclusions of this work.

@@

\section{Discussion and Conclusion}

@@

In this work, we aimed to analyse whether amphoric stamps could play an important role in the organisation of the workshops along rivers. For this reason, dissimilarity index was used to detect differences among workshops and stamps.

In our analysis, no strong relation between stamps and spatial distance have been detected in amphora workshops. The analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamp. Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found. In any case, our results show that most similar stamps were detected in the same \textit{conventus} area. These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.

The hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops. Rather, it seems that each workshop were organised independently with different stamps. Those stamps detected in closer workshops do not move from other distant workshops. In other words, the stamps tend to keep in the same area and different stamps were located in a same amphora workshops.

@@

>>>>

The paper addresses these questions as follows: the next section introduces the historical context. Section two displays the dataset and the methods used for the analysis. Section three presents the results and the last section shows the discussion and the main conclusions of this work.

@@

\section{Discussion}

@@

In this work, we aimed to analyse the distribution of the amphoric stamps in the organisation of the olive oil market both in the production and consumption areas. For this reason, a index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres. The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market.

\subsection{Production centres}

The analysis of the amphora workshops in Baetica province resulted a significance correlation between spatial distance and the similarity of the amphoric stamps.

The analysis suggests that there is  a connection between the same stamp code and the nearby amphora workshops, excluding certain exceptions. Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found. Indeed, our results show that most similar stamps were detected in the same \textit{conventus} area. This results can indicated a grouping of stamps with similar codes in the same \textit{conventus}. However,these stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and connection areas. In other words, we do not identify groups of more than three workshops or less sharing the same amphoric stamps as the dendrogram showed. In general, the majority of stamps were located in different amphora workshops.

Our results indicate that the hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops. Rather, it seems that each workshop were organised independently with different stamps. Those stamps detected in closer workshops do not move from other distant workshops. In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshops.

@@

\subsection{Consumption centres}

Both consumption centres showed a correlation between spatial distance and similarity in the amphoric stamps. In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}.

In \textit{Britannia}, the majority of similarity stamps were mainly found in military centres. This could be also interpreted by a intense bias where military centres have been mostly excavated than civil centres.

It is worth to mention that centres with a similarity correlation between  spatial distance and stamps were located close to the Atlantic coast around the North Sea and the Celtic Sea. This could indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.

This fact seems to concur with the latest research that reinforce the idea of a more important Atlantic route that would connect the provinces militarized.

(Remesal Rodrı́guez, 1986; 2008; Carreras Monfort y Morais, 2012; Morillo
et al., 2016; Rubio-Campillo et al., 2018b).

\maria{falta germania}

\section{Conclusion remarks}
